---
title: "Check - Full Shell Data"
author: "elisabethburnor"
date: "6/4/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}

library(lme4)
library(ggeffects)
library(splines)
library(ggplot2)
library(dplyr)
library(tidyr)
library(lme4)
library(emmeans)
library(splines)
library(stats)
library(piecewiseSEM)
library(kableExtra)
library(HLMdiag)
library(nlme)

```

```{r}

df <- readRDS("path_to_data_file.RDS")

```

```{r, echo = FALSE}

# Put data into long format
fs_df_long <- reshape(df, varying = list(c("dtifs_fa_fxrh", "dtifs_fa_fxlh"),
                                         c("dtifs_fa_cgcrh", "dtifs_fa_cgclh"),
                                         c("dtifs_fa_cghrh", "dtifs_fa_cghlh"),
                                         c("dtifs_fa_cstrh", "dtifs_fa_cstlh"),
                                         c("dtifs_fa_atrrh", "dtifs_fa_atrlh"),
                                         c("dtifs_fa_uncrh", "dtifs_fa_unclh"),
                                         c("dtifs_fa_ilfrh", "dtifs_fa_ilflh"),
                                         c("dtifs_fa_iforh", "dtifs_fa_ifolh"),
                                         c("dtifs_fa_slfrh", "dtifs_fa_slflh"),
                                         c("dtifs_fa_allfibrh", "dtifs_fa_allfiblh"),
                                         
                                         c("dtifs_md_fxrh", "dtifs_md_fxlh"),
                                         c("dtifs_md_cgcrh", "dtifs_md_cgclh"),
                                         c("dtifs_md_cghrh", "dtifs_md_cghlh"),
                                         c("dtifs_md_cstrh", "dtifs_md_cstlh"),
                                         c("dtifs_md_atrrh", "dtifs_md_atrlh"),
                                         c("dtifs_md_uncrh", "dtifs_md_unclh"),
                                         c("dtifs_md_ilfrh", "dtifs_md_ilflh"),
                                         c("dtifs_md_iforh", "dtifs_md_ifolh"),
                                         c("dtifs_md_slfrh", "dtifs_md_slflh"),
                                         c("dtifs_md_allfibrh", "dtifs_md_allfiblh"),
                                         
                                         c('dmri_rsin0_fiberat_cgcrh','dmri_rsin0_fiberat_cgclh'),
                                   c('dmri_rsin0_fiberat_cghrh','dmri_rsin0_fiberat_cghlh'),
                                   c('dmri_rsin0_fiberat_atrrh','dmri_rsin0_fiberat_atrlh'),
                                   c('dmri_rsin0_fiberat_uncrh','dmri_rsin0_fiberat_unclh'),
                                   c('dmri_rsin0_fiberat_iforh','dmri_rsin0_fiberat_ifolh'),
                                   c('dmri_rsin0_fiberat_slfrh','dmri_rsin0_fiberat_slflh'),
                                   c('dmri_rsin0_fiberat_fxrh','dmri_rsin0_fiberat_fxlh'),
                                   c('dmri_rsin0_fiberat_cstrh','dmri_rsin0_fiberat_cstlh'),
                                   c('dmri_rsin0_fiberat_ilfrh','dmri_rsin0_fiberat_ilflh'),
                                   c('dmri_rsin0_fiberat_allfibrh','dmri_rsin0_fiberat_allfiblh'),
                                   
                                   c('dmri_rsind_fiberat_cgcrh','dmri_rsind_fiberat_cgclh'),
                                   c('dmri_rsind_fiberat_cghrh','dmri_rsind_fiberat_cghlh'),
                                   c('dmri_rsind_fiberat_atrrh','dmri_rsind_fiberat_atrlh'),
                                   c('dmri_rsind_fiberat_uncrh','dmri_rsind_fiberat_unclh'),
                                   c('dmri_rsind_fiberat_iforh','dmri_rsind_fiberat_ifolh'),
                                   c('dmri_rsind_fiberat_slfrh','dmri_rsind_fiberat_slflh'),
                                   c('dmri_rsind_fiberat_fxrh','dmri_rsind_fiberat_fxlh'),
                                   c('dmri_rsind_fiberat_cstrh','dmri_rsind_fiberat_cstlh'),
                                   c('dmri_rsind_fiberat_ilfrh','dmri_rsind_fiberat_ilflh'),
                                   c('dmri_rsind_fiberat_allfibrh','dmri_rsind_fiberat_allfiblh'),
                                   
                                   c('dmri_rsints2_fiberat_cgcrh','dmri_rsints2_fiberat_cgclh'),
                                   c('dmri_rsints2_fiberat_cghrh','dmri_rsints2_fiberat_cghlh'),
                                   c('dmri_rsints2_fiberat_atrrh','dmri_rsints2_fiberat_atrlh'),
                                   c('dmri_rsints2_fiberat_uncrh','dmri_rsints2_fiberat_unclh'),
                                   c('dmri_rsints2_fiberat_iforh','dmri_rsints2_fiberat_ifolh'),
                                   c('dmri_rsints2_fiberat_slfrh','dmri_rsints2_fiberat_slflh'),
                                   c('dmri_rsints2_fiberat_fxrh','dmri_rsints2_fiberat_fxlh'),
                                   c('dmri_rsints2_fiberat_cstrh','dmri_rsints2_fiberat_cstlh'),
                                   c('dmri_rsints2_fiberat_ilfrh','dmri_rsints2_fiberat_ilflh'),
                                   c('dmri_rsints2_fiberat_allfibrh','dmri_rsints2_fiberat_allfiblh')),
                      
                      v.names = c("fs_fa_fx", "fs_fa_cgc", "fs_fa_cgh", "fs_fa_cst", "fs_fa_atr",
                                  "fs_fa_unc", "fs_fa_ilf", "fs_fa_ifo", "fs_fa_slf", "fs_fa_allfib",
                                  
                                  "fs_md_fx", "fs_md_cgc", "fs_md_cgh", "fs_md_cst", "fs_md_atr",
                                  "fs_md_unc", "fs_md_ilf", "fs_md_ifo", "fs_md_slf", "fs_md_allfib",
                                  
                                  'rsin0_cgc', 'rsin0_cgh', 'rsin0_atr', 'rsin0_unc', 'rsin0_ifo', 'rsin0_slf',
                                  'rsin0_fx', 'rsin0_cst', 'rsin0_ilf', 'rsin0_allfib',
                            
                            'rsind_cgc', 'rsind_cgh', 'rsind_atr', 'rsind_unc', 'rsind_ifo', 'rsind_slf',
                            'rsind_fx', 'rsind_cst', 'rsind_ilf', 'rsind_allfib',
                            
                            'rsints2_cgc', 'rsints2_cgh', 'rsints2_atr', 'rsints2_unc', 'rsints2_ifo',
                            'rsints2_slf', 'rsints2_fx', 'rsints2_cst', 'rsints2_ilf', 'rsints2_allfib'), 
                      idvar = "subjectkey", timevar= "hemisphere", times = c("right", "left"), direction = "long")

```


```{r outcome setup, echo = FALSE}
# Data

savepath <- "path_for_output"
my.knots <- attr(ns(df$pm2.5, 3), "knots")

diffusion = c(rep("DTI", 20), rep("RSI", 30))

dti = c(rep("FA", 10), rep("MD", 10), rep("N0", 10), rep("ND", 10), rep("NTS2", 10))

outcomes <- c("fs_fa_cgc", "fs_fa_cgh", "fs_fa_atr", "fs_fa_unc", "fs_fa_ifo", "fs_fa_slf", "fs_fa_fx", "fs_fa_cst",
              "fs_fa_ilf", "fs_fa_allfib",
              "fs_md_cgc", "fs_md_cgh", "fs_md_atr", "fs_md_unc", "fs_md_ifo", "fs_md_slf", "fs_md_fx","fs_md_cst",
              "fs_md_ilf", "fs_md_allfib", 
              
              'rsin0_cgc', 'rsin0_cgh', 'rsin0_atr', 'rsin0_unc', 'rsin0_ifo', 'rsin0_slf', 
              'rsin0_fx', 'rsin0_cst', 'rsin0_ilf', 'rsin0_allfib',
              'rsind_cgc', 'rsind_cgh', 'rsind_atr', 'rsind_unc', 'rsind_ifo', 'rsind_slf',
              'rsind_fx', 'rsind_cst', 'rsind_ilf', 'rsind_allfib',
              'rsints2_cgc', 'rsints2_cgh', 'rsints2_atr', 'rsints2_unc', 'rsints2_ifo',
              'rsints2_slf', 'rsints2_fx', 'rsints2_cst', 'rsints2_ilf', 'rsints2_allfib')

tracts <- rep(c("cgc", "cgh", "atr", "unc", "ifo", "slf", "fx", "cst", "ilf", "allfib"), 5)
tract_names <- toupper(all_tracts)

long_key <- data.frame(diffusion, dti, outcomes, tracts, tract_names)

```

```{r Fit interactions, echo = FALSE}

key <- long_key
df_HEI <- fs_df_long

interaction_anova_tbl <- data.frame()
linear_intx_anova_tbl <- data.frame()
null_anova_tbl <- data.frame()
nonlin_anova_tbl <- data.frame()

models <- list()
for (i in 1:length(outcomes)) {

  tract <- key$tract_names[which(key$outcomes == outcomes[i])]
  dti <- key$dti[which(key$tract_names == tract)]
  
  # Fit model with interaction term for hemisphere
  mo_hem_intx <- lmer(as.formula(paste(outcomes[i], "~", 'ns(pm2.5, knots = my.knots)*hemisphere + sex + interview_age + race_ethnicity + high.educ + overall.income + demo_prnt_empl_v2 + neighb_phenx_avg_p + dmri_rsi_meanmotion + handedness + mri_info_manufacturer + (1 | abcd_site) + (1 | abcd_site:subjectkey)')), data = df_HEI, REML = FALSE, na.action = na.omit)
  
  # Fit model without interaction term for hemisphere
  mo_hem <- lmer(as.formula(paste(outcomes[i], "~", 'ns(pm2.5, 3) + hemisphere + sex + interview_age + race_ethnicity + high.educ + overall.income + demo_prnt_empl_v2 + neighb_phenx_avg_p + dmri_rsi_meanmotion + handedness + mri_info_manufacturer + (1 | abcd_site) + (1 | abcd_site:subjectkey)')), data = df_HEI, REML = FALSE, na.action = na.omit)

  check_intx <- anova(mo_hem, mo_hem_intx)

  intx_anova <- data.frame("Outcome" = outcomes[i], "Tract" = tract, "p-value" = check_intx[2,"Pr(>Chisq)"])
  interaction_anova_tbl <- rbind(interaction_anova_tbl, intx_anova)

  null_mo <- lmer(as.formula(paste(outcomes[i], "~", 'hemisphere + sex + interview_age + race_ethnicity + high.educ + overall.income + demo_prnt_empl_v2 + neighb_phenx_avg_p + dmri_rsi_meanmotion + handedness + mri_info_manufacturer + (1 | abcd_site) + (1 | abcd_site:subjectkey)')),
                      data = df_HEI, REML = FALSE, na.action = na.omit)
  
  linear_mo <- lmer(as.formula(paste(outcomes[i], "~", 'pm2.5 + hemisphere + sex + interview_age + race_ethnicity + high.educ + overall.income + demo_prnt_empl_v2 + neighb_phenx_avg_p + dmri_rsi_meanmotion + handedness + mri_info_manufacturer + (1 | abcd_site) + (1 | abcd_site:subjectkey)')), 
                      data = df_HEI, REML = FALSE, na.action = na.omit)
  
  linear_mo_intx <- lmer(as.formula(paste(outcomes[i], "~", 'pm2.5*hemisphere + sex + interview_age + race_ethnicity + high.educ + overall.income + demo_prnt_empl_v2 + neighb_phenx_avg_p + dmri_rsi_meanmotion + handedness + mri_info_manufacturer + (1 | abcd_site) + (1 | abcd_site:subjectkey)')), 
                      data = df_HEI, REML = FALSE, na.action = na.omit)
  
  check_lin_intx <- anova(linear_mo, linear_mo_intx)
  lin_intx_anova <- data.frame("Tract" = tract, "p_value" = check_lin_intx$`Pr(>Chisq)`[2])
  linear_intx_anova_tbl <- rbind(linear_intx_anova_tbl, lin_intx_anova)
      
  check_null <- anova(null_mo, mo_hem_intx)
  null_anova <- data.frame("Outcome" = outcomes[i], "Tract" = tract, "p-value" = check_null[2,"Pr(>Chisq)"])
  null_anova_tbl <- rbind(null_anova_tbl, null_anova)

  linear_anova <- anova(linear_mo_intx, mo_hem_intx)
  nonlin_anova <- data.frame("Outcome" = outcomes[i], "Tract" = tract, "p-value" = linear_anova[2,"Pr(>Chisq)"])
  nonlin_anova_tbl <- rbind(nonlin_anova_tbl, nonlin_anova)
  
  r2 <- rsquared(mo_hem_intx)

  models[[tract]][['spline_model']] <- mo_hem_intx
  models[[tract]][['linear_model']] <- linear_mo_intx
  models[[tract]][['analyses']]$null_anova <- null_anova
  models[[tract]][['analyses']]$linear_anova <- linear_anova
  models[[tract]][['analyses']]$interaction_check <- intx_anova
  models[[tract]][['analyses']]$r2 <- r2
  models[[tract]][['outcome']] <- outcomes[i]
}

```

# Print null anovas
```{r Print null anova tables}

# Null Table
null_anova_tbl$adj_p_value <- p.adjust(null_anova_tbl$p.value, method = "BH")

dti_label <- function(outcome) {
  label <- long_key[which(long_key$outcomes == outcome),"dti"]
  return(label)
}

null_anova_tbl$dti <- sapply(null_anova_tbl$Outcome, dti_label)
null_anova_tbl <- null_anova_tbl %>% relocate(dti, .after = Outcome)


kbl(null_anova_tbl, caption = "ANOVA comparison: Spline model with Hemisphere Interaction vs. Null model") %>%
  kable_classic(full_width = TRUE)

# list of non-null models

sig_non_null <- c()
for (i in 1:nrow(null_anova_tbl)) {
  if (null_anova_tbl$adj_p_value[i] < 0.05) sig_non_null <- c(sig_non_null, null_anova_tbl$Outcome[i])
}

```

# Check residuals
```{r, include = FALSE}

resids <- function(model_list, anova_results) {
  
  sig_mods <- which(anova_results$adj_p_value < 0.05)
  sig_tracts <- anova_results$Tract[sig_mods]
  
  for (j in 1:length(sig_tracts)) {
    
    tract <- sig_tracts[[j]]
    i <- which(key$tract_names == tract)
    outcome <- key$outcomes[i]
    
    formula <- paste0(outcome, "~", 'ns(pm2.5, 3)*hemisphere + sex + interview_age + race_ethnicity + high.educ + overall.income + demo_prnt_empl_v2 + neighb_phenx_avg_p + dmri_rsi_meanmotion + handedness + mri_info_manufacturer + (1 | abcd_site) + (1 | abcd_site:subjectkey)')
    
    mo <- lmer(as.formula(formula), data = df_HEI, REML = FALSE, na.action = na.omit)

    
    lvl1_res <- hlm_resid(mo, level = 1, standardize = TRUE, include.ls = FALSE)
    res_vs_fitted <- qplot(x = .fitted, y = .std.resid, 
                             data = lvl1_res, geom = c("point", "smooth"),
                            main = paste0(model_list$name, " ", tract, 
                                          "\nLevel-1 Std. LS Residuals vs. Fitted Values"),
                            ylab= "Std. LS Level-1 Residuals", xlab = "Fitted Values")
    print(res_vs_fitted)
    qqnorm(lvl1_res$.std.resid, main = paste0(model_list$name, " ", tract, 
                                          "\nLevel 1 Std. LS Residuals: Normal Q-Q Plot"))
    qqline(lvl1_res$.std.resid, col = 2, lwd=2,lty=2)
  }
}

```

# Anovas to check linearity (splines vs. linear)
```{r Print linear anova tables}

# Non-Linearity table

nonlin_anova_tbl$dti <- sapply(nonlin_anova_tbl$Outcome, dti_label)
nonlin_anova_tbl <- nonlin_anova_tbl %>% relocate(dti, .after = Outcome)

nonlin_anova_tbl$adj_p_value <- p.adjust(nonlin_anova_tbl$p.value, method = "BH")
kbl(nonlin_anova_tbl, caption = "ANOVA comparison: Spline model with interaction term vs. Linear model with interaction term") %>%
  kable_classic(full_width = TRUE)

```

# Anovas to check hemisphere interaction
```{r hem intx table}
# Hemisphere interaction table

interaction_anova_tbl$adj_p_value <- p.adjust(interaction_anova_tbl$p.value, method = "BH")

interaction_anova_tbl$dti <- sapply(interaction_anova_tbl$Outcome, dti_label)
interaction_anova_tbl <- interaction_anova_tbl %>% relocate(dti, .after = Outcome)

kbl(interaction_anova_tbl, caption = "ANOVA comparison: Hemisphere Interaction vs. Hemipshere as fixed effect") %>%
  kable_classic(full_width = TRUE)

# Make a list of models with significant interaction terms for post-hoc analysis

sig_intx <- c()
for (i in 1:length(sig_non_null)) {
  k <- which(interaction_anova_tbl$Outcome == sig_non_null[i])
  if (interaction_anova_tbl$adj_p[k] < 0.05) sig_intx <- c(sig_intx, interaction_anova_tbl$Outcome[k])
}

interactions <- data.frame()
for (i in 1:length(sig_intx)) {
  row <- interaction_anova_tbl[which(interaction_anova_tbl$Outcome == sig_intx[i]),]
  interactions <- rbind(interactions, row)
}
row.names(interactions) <- NULL
kbl(interactions, digits = 4,
    caption = "Models with Significant Interaction Terms (adj. p < 0.05)") %>%
    kable_classic(full_width = FALSE)

#write.csv(interaction_anova_tbl, 
#          paste0(savepath, diffusion, "_", measure, "_intx_anova_tbl.csv"))

```

# R-squared values
```{r r-squared}

r2_df <- data.frame()
for (i in 1:length(models)) {
  r2_df <- rbind(r2_df, models[[i]]$analyses$r2)
}

r2_df <- cbind(r2_df$Response, round(r2_df$Marginal, 4), round(r2_df$Conditional, 4))
colnames(r2_df) <- c("Model", "Marginal", "Conditional")

kbl(r2_df, caption = "R-squared Values") %>%
  kable_classic(full_width = TRUE)

write.csv(r2_df, 
          paste0(savepath, "md_r2values.csv"))
```

## Wide analysis
```{r}
# Wide dataset key

wide_outcomes <- c("dtifs_fa_cgcrh", "dtifs_fa_cgclh", "dtifs_fa_cghrh", "dtifs_fa_cghrh", "dtifs_fa_atrrh", "dtifs_fa_atrlh",
                   "dtifs_fa_uncrh", "dtifs_fa_unclh", "dtifs_fa_iforh", "dtifs_fa_ifolh", "dtifs_fa_slfrh", "dtifs_fa_slflh",
                   "dtifs_fa_fxrh", "dtifs_fa_fxlh", "dtifs_fa_cstrh", "dtifs_fa_cstlh", "dtifs_fa_ilfrh", "dtifs_fa_ilflh",
                   "dtifs_fa_allfibrh", "dtifs_fa_allfiblh", "dtifs_md_cc",
                   
                   "dtifs_md_cgcrh", "dtifs_md_cgclh", "dtifs_md_cghrh", "dtifs_md_cghlh", "dtifs_md_atrrh", "dtifs_md_atrlh",
                   "dtifs_md_uncrh", "dtifs_md_unclh", "dtifs_md_iforh", "dtifs_md_ifolh", "dtifs_md_slfrh", "dtifs_md_slflh",
                   "dtifs_md_fxrh", "dtifs_md_fxlh", "dtifs_md_cstrh", "dtifs_md_cstlh", "dtifs_md_ilfrh", "dtifs_md_ilflh",
                   "dtifs_md_allfibrh", "dtifs_md_allfiblh", "dtifs_fa_cc",
              
                                  'dmri_rsin0_fiberat_cgcrh','dmri_rsin0_fiberat_cgclh',
                                  'dmri_rsin0_fiberat_cghrh','dmri_rsin0_fiberat_cghlh',
                                  'dmri_rsin0_fiberat_atrrh','dmri_rsin0_fiberat_atrlh',
                                  'dmri_rsin0_fiberat_uncrh','dmri_rsin0_fiberat_unclh',
                                   'dmri_rsin0_fiberat_iforh','dmri_rsin0_fiberat_ifolh',
                                   'dmri_rsin0_fiberat_slfrh','dmri_rsin0_fiberat_slflh',
                                   'dmri_rsin0_fiberat_fxrh','dmri_rsin0_fiberat_fxlh',
                                   'dmri_rsin0_fiberat_cstrh','dmri_rsin0_fiberat_cstlh',
                                   'dmri_rsin0_fiberat_ilfrh','dmri_rsin0_fiberat_ilflh',
                                   'dmri_rsin0_fiberat_allfibrh','dmri_rsin0_fiberat_allfiblh', 'dmri_rsin0_fiberat_cc',
                                   
                                   'dmri_rsind_fiberat_cgcrh','dmri_rsind_fiberat_cgclh',
                                   'dmri_rsind_fiberat_cghrh','dmri_rsind_fiberat_cghlh',
                                   'dmri_rsind_fiberat_atrrh','dmri_rsind_fiberat_atrlh',
                                   'dmri_rsind_fiberat_uncrh','dmri_rsind_fiberat_unclh',
                                   'dmri_rsind_fiberat_iforh','dmri_rsind_fiberat_ifolh',
                                   'dmri_rsind_fiberat_slfrh','dmri_rsind_fiberat_slflh',
                                   'dmri_rsind_fiberat_fxrh','dmri_rsind_fiberat_fxlh',
                                   'dmri_rsind_fiberat_cstrh','dmri_rsind_fiberat_cstlh',
                                   'dmri_rsind_fiberat_ilfrh','dmri_rsind_fiberat_ilflh',
                                   'dmri_rsind_fiberat_allfibrh','dmri_rsind_fiberat_allfiblh', 'dmri_rsind_fiberat_cc',
                                   
                                   'dmri_rsints2_fiberat_cgcrh','dmri_rsints2_fiberat_cgclh',
                                   'dmri_rsints2_fiberat_cghrh','dmri_rsints2_fiberat_cghlh',
                                   'dmri_rsints2_fiberat_atrrh','dmri_rsints2_fiberat_atrlh',
                                   'dmri_rsints2_fiberat_uncrh','dmri_rsints2_fiberat_unclh',
                                   'dmri_rsints2_fiberat_iforh','dmri_rsints2_fiberat_ifolh',
                                   'dmri_rsints2_fiberat_slfrh','dmri_rsints2_fiberat_slflh',
                                   'dmri_rsints2_fiberat_fxrh','dmri_rsints2_fiberat_fxlh',
                                   'dmri_rsints2_fiberat_cstrh','dmri_rsints2_fiberat_cstlh',
                                   'dmri_rsints2_fiberat_ilfrh','dmri_rsints2_fiberat_ilflh',
                                   'dmri_rsints2_fiberat_allfibrh','dmri_rsints2_fiberat_allfiblh', 'dmri_rsints2_fiberat_cc')

hemisphere = c(rep(c(rep(c("Right", "Left"), 10), "-"), 5))

tracts <- rep(c("cgc", "cgc", "cgh", "cgh", "atr", "atr", "unc", "unc", "ifo", "ifo", "slf", "slf", "fx", "fx", "cst", 
              "cst", "ilf", "ilf", "allfib", "allfib", "cc"), 5)

tract_names = toupper(tracts)

key_sep <- data.frame(outcomes = wide_outcomes, hemisphere = hemisphere, tracts = tracts, tract_names = tract_names)

```

```{r, hemisphere-strat models, echo = FALSE}
# Re-fit models by hemisphere

df_HEI_wide = df
sep_models <- list()

my.knots <- attr(ns(df$pm2.5, 3), "knots")

sep_null_anova <- data.frame()
sep_lin_anova <- data.frame()

for (i in 1:nrow(key_sep)) {
  
  tract <- key_sep$tracts[i]
  
  ### Left
  outcome <- key_sep$outcomes[i]
  hemisphere <- key_sep$hemisphere[i]
  
  crude <- lmer(as.formula(paste(outcome, "~", 'ns(pm2.5, knots = my.knots) + sex + interview_age  + (1 | abcd_site)')),  data = df_HEI_wide, REML = FALSE, na.action = na.omit)
  
  mo <-  lmer(as.formula(paste(outcome, "~", 'ns(pm2.5, knots = my.knots) + sex + interview_age + race_ethnicity + high.educ + overall.income + demo_prnt_empl_v2 + neighb_phenx_avg_p + dmri_rsi_meanmotion + handedness + mri_info_manufacturer + (1 | abcd_site)')), 
                   data = df_HEI_wide, REML = FALSE, na.action = na.omit)
  
  null_mo <- lmer(as.formula(paste(outcome, "~", 'sex + interview_age + race_ethnicity + high.educ + overall.income + demo_prnt_empl_v2 + neighb_phenx_avg_p + dmri_rsi_meanmotion + handedness + mri_info_manufacturer + (1 | abcd_site)')), 
                   data = df_HEI_wide, REML = FALSE, na.action = na.omit)
  
  lin_mo <- lmer(as.formula(paste(outcome, "~", 'pm2.5 + sex + interview_age + race_ethnicity + high.educ + overall.income + demo_prnt_empl_v2 + neighb_phenx_avg_p + dmri_rsi_meanmotion + handedness + mri_info_manufacturer + (1 | abcd_site)')), 
                   data = df_HEI_wide, REML = FALSE, na.action = na.omit)
  
  check_null <- anova(null_mo, mo)
  null_df <- data.frame("Outcome" = outcome, "Tract" = tract, "Hemisphere" = hemisphere, "p-value" = check_null[2, "Pr(>Chisq)"])
  sep_null_anova <- rbind(sep_null_anova, null_df)
  
  check_lin <- anova(lin_mo, mo)
  lin_df <- data.frame("Outcome" = outcome, "Tract" = tract, "Hemisphere" = hemisphere, "p-value" = check_lin[2, "Pr(>Chisq)"])
  sep_lin_anova <- rbind(sep_lin_anova, lin_df)
  
  sep_models[[outcome]]$spline_mod <- mo
  sep_models[[outcome]]$lin_mod <- lin_mo
  sep_models[[outcome]]$crude <- crude
  sep_models[[outcome]]$tract <- tract
  sep_models[[outcome]]$outcome <- outcome
  sep_models[[outcome]]$hemisphere <- hemisphere
  sep_models[[outcome]]$r2 <- rsquared(mo)
}

main_analysis_sep_models = sep_models
```

# R-squared for hemisphere-stratified models
```{r r-squared separate models}

r2_df_sep <- data.frame()
for (i in 1:length(sep_models)) {
    outcome <- sep_models[[i]]$outcome
    tract <- sep_models[[i]]$tract
    hemisphere <- sep_models[[i]]$hemisphere
    row <- cbind(tract, hemisphere, sep_models[[i]]$r2)
    r2_df_sep <- rbind(r2_df_sep, row)
}

r2_df_sep <- as.data.frame(cbind(r2_df_sep$tract, r2_df_sep$hemisphere, r2_df_sep$Response, 
                   round(r2_df_sep$Marginal, 4), round(r2_df_sep$Conditional, 4)))
colnames(r2_df_sep) <- c("Tract", "Hemisphere", "Model", "Marginal", "Conditional")

kbl(r2_df_sep, caption = "R-squared Values") %>%
  kable_classic(full_width = TRUE)

```

# Print tables for non-null hemisphere-stratified models
```{r}

dti_label_sep <- function(outcome) {
  label <- key_sep[which(key_sep$outcomes == outcome),"dti"]
  return(label)
}

non_null_models <- sep_null_anova[which(sep_null_anova$p.value < 0.05),]
if (nrow(non_null_models)==0) {
  print("No non-null models")
} else {
  sep_hem_models <- data.frame()
  
  for (i in 1:nrow(non_null_models)) {
    outcome <- non_null_models$Outcome[i]
    tract <- non_null_models$Tract[i]
    hemisphere <- tolower(non_null_models$Hemisphere[i])
    
    row <- which(r2_df_sep$Model == outcome)
    
    newrow <- cbind(non_null_models[i,], r2_df_sep[row,"Marginal"], r2_df_sep[row,"Conditional"])
    sep_hem_models <- rbind(sep_hem_models, newrow)
  }
  

  
  sep_hem_models$dti <- sapply(sep_hem_models$Outcome, dti_label_sep)
  sep_hem_models <- sep_hem_models %>% relocate(dti, .after = Outcome)
  
  
  colnames(sep_hem_models) <- c("Outcome", "DTI", "Tract", "Hemisphere", "p_value", "Marginal", "Conditional")
  
  row.names(sep_hem_models) <- NULL
  kbl(sep_hem_models, digits = 4, 
      caption = "Hemisphere-Stratified Models with Significant Fits (non-Null, p < 0.05)") %>%
      kable_classic(full_width = TRUE)
}

```

## Plot significant models
```{r, warning = FALSE, message = FALSE}

tracts <- c("unc", "atr", "slf", "cgh", "ifo", "cc", "allfib")

md_sig_models = names(sep_models)[c(22, 23, 24, 25, 26, 30, 32,34, 40)]
# MD colors
tracts = c("atr", "cgh", "ifo", "slf", "unc", "cc", "allfib")
color_strings = c("#00FF00", "yellow", "olivedrab3", "cyan3", "skyblue", "red", "black")
cname = c("bright green", "yellow", "olive", "cyan", "blue", "red", "black")
color_key <- data.frame(tracts, cname, color_strings)

for (i in 1:length(cc_models)) {
    
  model = sep_models[[which(names(sep_models) == md_sig_models[i])]]
  
  outcome = model$outcome
  hem = str_to_title(model$hemisphere)
  tract = model$tract
  mo = model$spline_mod
  df = df_HEI_wide
  
  row <- which(color_key$tracts == tract)
  plot_color <- color_key$color_strings[row]

  grid <- ref_grid(mo, at = list(pm2.5 = seq(3, 13, by = 0.5), age = c(112, 116)), data = df, params = "my.knots")
    
  emm <- as.data.frame(emmeans(grid, "pm2.5", params = "my.knots"))
    
  colnames(emm) <- c("pm2.5", "emmean", "SE", "df", "LCL", "UCL")
    
  plot <- ggplot(data = emm, aes(x = pm2.5)) +
              geom_line(aes(y = emmean), color = plot_color, size = .8) +
              geom_ribbon(aes(ymin = LCL, ymax = UCL), fill = plot_color, alpha = 0.3) +
              xlim(c(3, 13)) +
              geom_rug(data = df,
                       aes(x = pm2.5),
                       alpha = 0.2) + 
              scale_linetype_manual() +
              xlab(expression(PM[2.5]~(ug/m^{"3"}))) +
              ylab(paste0("Mean Diffusivity (MD)")) +
              ggtitle(label = paste0(toupper(tract), " - ", hem, " Hemisphere")) + 
              theme(panel.grid.major = element_line(colour = "grey95"), 
                    panel.grid.minor = element_blank(),
                    panel.background = element_blank(), 
                    title = element_text(size = 36),
                    axis.line = element_line(colour = "black"),
                    axis.title.x = element_text(size = 32),
                    axis.title.y = element_text(size = 32),
                    axis.text = element_text(size = 32),
                    strip.text = element_text(size = 32))
    print(plot)
    
    
    filename <- paste0(outcome, '_sensitivity.tiff')
    tiff(paste0(savepath, filename),
               height = 20, width = 25, units='cm',compression = "lzw", res = 300)
    print(plot)
    dev.off()
}
 
tracts <- c("unc", "atr", "slf", "cgh", "ifo", "cc", "allfib")

# md_sig_models = names(sep_models)[c(22, 23, 24, 25, 26, 30, 32,34, 40)]
# # MD colors
# tracts = c("atr", "cgh", "ifo", "slf", "unc", "cc", "allfib")
# color_strings = c("#00FF00", "yellow", "olivedrab3", "cyan3", "skyblue", "red", "black")
# cname = c("bright green", "yellow", "olive", "cyan", "blue", "red", "black")
# color_key <- data.frame(tracts, cname, color_strings)
# 
# ggplot(data = df_HEI_wide, aes(x = pm2.5, y = interview_age)) +
#   geom_point(color = "skyblue")

  

```

```{r Functions to calculate percent change, echo = FALSE}

# Percent Change Helper Functions
# Factor variable levels

set_levels <- function(mo, df) {

  high_education <- levels(df$high.educ)
  income <- levels(df$overall.income)
  income <- income[1:3]
  age_level_6 <- as.integer(c(119, 125))
  
  # PM2.5
  pmlow <- c(4,8)
  pmhigh <- c(8,12)
  pm25_est <- as.data.frame(emmeans(mo, "pm2.5", 
                                   at = list(pm2.5 = c(4, 8, 12)), 
                                   data = df, params = "my.knots"))
  pm25_est$cov <- "pm2.5"
  colnames(pm25_est) <- c("cov_level", "emmean", "SE", "df", "lower.CL", "upper.CL", "covariate")


  # Age
  age_est <- as.data.frame(emmeans(mo, "interview_age", 
                                   at = list(interview_age = age_level_6), data = df, params = "my.knots"))
  age_est$covariate = "age"
  colnames(age_est) <- c("cov_level", "emmean", "SE", "df", "lower.CL", "upper.CL", "covariate")

  # Parents' education
  educ_est <- as.data.frame(emmeans(mo, "high.educ",
                                    at = list(high.educ = high_education), data = df, params = "my.knots"))
  educ_est$covariate <- "parent_educ"
  colnames(educ_est) <- c("cov_level", "emmean", "SE", "df", "lower.CL", "upper.CL", "covariate")

  # Household income
  income_est <- as.data.frame(emmeans(mo, "overall.income",
                                    at = list(overall.income = income), data = df, params = "my.knots"))
  income_est$covariate = "income"
  colnames(income_est) <- c("cov_level", "emmean", "SE", "df", "lower.CL", "upper.CL", "covariate")

  est_table <- rbind(pm25_est, age_est, educ_est, income_est)
  return(est_table)
}

make_pc_table <- function(model, data, outcome, hemisphere, tract){
  quantile_est <- set_levels(model, data)
  variables <- unique(quantile_est[c("covariate")])
  perc_change_table <- data.frame()

  for (i in 1:nrow(variables)) {
    var <- filter(quantile_est, covariate == variables$covariate[i])
    j = 1
    pc_df <- data.frame()
    while (j <= (nrow(var) - 1)) {
      percent_change_est <- perc_change(var[j, 2], var[j, 3], var[j+1, 2], var[j+1, 3])
      levels <- paste0(var[j, 1], " -> ", var[j+1, 1])
      pc_df <- rbind(pc_df, cbind(var$covariate[1], levels, percent_change_est))
      j <- j + 1
    }
    perc_change_table <- rbind(perc_change_table, pc_df)
  }
  
  rounded_pc_table <- cbind(perc_change_table[,1:2], round(perc_change_table[,3:6], digits = 3))
  rounded_pc_table <- cbind(outcome, tract, hemisphere,
                            rounded_pc_table[,1:4], paste0("(", rounded_pc_table[,5], ", ", rounded_pc_table[,6], ")"))
  colnames(rounded_pc_table) <- c("Outcome", "Tract", "Hemisphere", "Predictor", "Levels", "Percent Change", "SE", "95% CI")
  return(rounded_pc_table)
}

perc_change <- function(est_first, se_first, est_second, se_second)
{

  percent_change <- ((est_second - est_first) / est_first) * 100
  percent_change_se <- abs(est_second/est_first) * sqrt((se_second^2/est_second^2) +
                                                        (se_first^2/est_first^2)) * 100
  conf.High <- percent_change + 1.96*percent_change_se
  conf.Low <- percent_change - 1.96*percent_change_se
  return_df <- data.frame(percent_change, percent_change_se, conf.Low, conf.High)
  return(return_df)
}

```

# Print Percent Change tables
```{r, warning = FALSE, message = FALSE}

pc_summary <- data.frame()

for (i in 1:length(sep_models)) {
  mo <- sep_models[[i]]
  model <- mo$spline_mod
  tract <- mo$tract
  outcome <- mo$outcome
  hemisphere <- mo$hemisphere
  
  pc <- make_pc_table(model, df_HEI_wide, outcome, hemisphere, tract)
  pc_summary <- rbind(pc_summary, pc)
  
}

main_analysis_pc_summary = pc_summary


hem = c("left", "right")
for (j in 1:2) {
  hem_df = filter(pc_summary, Hemisphere == hem[j])
  for (i in 1:length(hei_tracts)) {
    tract_df = filter(hem_df, Tract == hei_tracts[i])
    tract_df = filter(tract_df, Predictor == "pm2.5")
    
    md = tract_df[grep("md", tract_df$Outcome),]
    fa = tract_df[grep("fa", tract_df$Outcome),]
    hei_fa_pc_table = rbind(hei_fa_pc_table, fa)
    hei_md_pc_table = rbind(hei_md_pc_table, md)
  }
}

hei_pc_summary = rbind(hei_fa_pc_table, hei_md_pc_table)

write.csv(hei_pc_summary, paste0(savepath, "dti_sensitivity_pc_tables.csv"))
```

# Sensitivity Analysis for subjects with PM2.5 <= 12
```{r}

df_HEI_wide = df
df_HEI_wide = filter(df_HEI_wide, pm2.5 <= 12)

sep_models <- list()

my.knots <- attr(ns(df$pm2.5, 3), "knots")

sep_null_anova <- data.frame()
sep_lin_anova <- data.frame()

for (i in 1:nrow(key_sep)) {
  
  tract <- key_sep$tracts[i]
  
  ### Left
  outcome <- key_sep$outcomes[i]
  hemisphere <- key_sep$hemisphere[i]
  
  crude <- lmer(as.formula(paste(outcome, "~", 'ns(pm2.5, knots = my.knots) + sex + interview_age  + (1 | abcd_site)')),  data = df_HEI_wide, REML = FALSE, na.action = na.omit)
  
  mo <-  lmer(as.formula(paste(outcome, "~", 'ns(pm2.5, knots = my.knots) + sex + interview_age + race_ethnicity + high.educ + overall.income + demo_prnt_empl_v2 + neighb_phenx_avg_p + dmri_rsi_meanmotion + handedness + mri_info_manufacturer + (1 | abcd_site)')), 
                   data = df_HEI_wide, REML = FALSE, na.action = na.omit)
  
  null_mo <- lmer(as.formula(paste(outcome, "~", 'sex + interview_age + race_ethnicity + high.educ + overall.income + demo_prnt_empl_v2 + neighb_phenx_avg_p + dmri_rsi_meanmotion + handedness + mri_info_manufacturer + (1 | abcd_site)')), 
                   data = df_HEI_wide, REML = FALSE, na.action = na.omit)
  
  lin_mo <- lmer(as.formula(paste(outcome, "~", 'pm2.5 + sex + interview_age + race_ethnicity + high.educ + overall.income + demo_prnt_empl_v2 + neighb_phenx_avg_p + dmri_rsi_meanmotion + handedness + mri_info_manufacturer + (1 | abcd_site)')), 
                   data = df_HEI_wide, REML = FALSE, na.action = na.omit)
  
  check_null <- anova(null_mo, mo)
  null_df <- data.frame("Outcome" = outcome, "Tract" = tract, "Hemisphere" = hemisphere, "p-value" = check_null[2, "Pr(>Chisq)"])
  sep_null_anova <- rbind(sep_null_anova, null_df)
  
  check_lin <- anova(lin_mo, mo)
  lin_df <- data.frame("Outcome" = outcome, "Tract" = tract, "Hemisphere" = hemisphere, "p-value" = check_lin[2, "Pr(>Chisq)"])
  sep_lin_anova <- rbind(sep_lin_anova, lin_df)
  
  sep_models[[outcome]]$spline_mod <- mo
  sep_models[[outcome]]$lin_mod <- lin_mo
  sep_models[[outcome]]$crude <- crude
  sep_models[[outcome]]$tract <- tract
  sep_models[[outcome]]$outcome <- outcome
  sep_models[[outcome]]$hemisphere <- hemisphere
  sep_models[[outcome]]$r2 <- rsquared(mo)
}

sensitivity_sep_models = sep_models

```

# Main analysis vs. sensitivity analysis comparison
```{r}

pc_summary <- data.frame()
sep_models = sensitivity_sep_models
for (i in 1:length(sep_models)) {
  mo <- sep_models[[i]]
  model <- mo$spline_mod
  tract <- mo$tract
  outcome <- mo$outcome
  hemisphere <- mo$hemisphere
  
  pc <- make_pc_table(model, df_HEI_wide, outcome, hemisphere, tract)
  pc_summary <- rbind(pc_summary, pc)
  
}

pc_summary_sensitivity = pc_summary

compare_pc = cbind(main_analysis_pc_summary, pc_summary_sensitivity)
```


